---
title: "Session 2"
author: ""
date: "11 12 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load the tidyverse
```{r}
pacman::p_load(tidyverse)
```

set a seed
```{r}
set.seed(2020)
```

Use a new code snippet for every exercise.

1. load the 2012 French presidential election data
```{r}
load("runoff.Rda") # pay attention where you have saved the runoff data set on your PC!
```

2. inspect the data
```{r}
glimpse(runoff)
```

3. add a unique voter id, the total number of voters and the number of voters by municipality
```{r}
runoff <- 
  runoff %>%
  mutate(voter_id = 1:nrow(.),
         turnout = nrow(.)) %>%
  left_join(runoff %>%
              group_by(municipality) %>%
              count(name = "turnout_municipality"))
```

4. inspect the data again (check if new variables are correct)
```{r}
glimpse(runoff)
```

5. create a variable sample_size
```{r}
sample_size <- 2500
```

6. draw a simple random sample
```{r}
srs <- 
  runoff %>% 
  sample_n(sample_size)

glimpse(srs)
```

7. draw a stratified sample (strata = municipality, sample size = sample_size, size of strata = adjusted to number of voters per municipality)
```{r}
str <- 
  runoff %>% 
  mutate(voters_strata = round(sample_size * turnout_municipality / turnout)) %>%
  group_by(municipality) %>%
  sample_n(voters_strata) %>%
  ungroup()

glimpse(str)
```

8. create a variable number_clusters
```{r}
number_clusters <- 5
```

9. draw a cluster sample with probability of being sampled according to the number of voters by municipality (clusters = municipality, number of clusters = number_clusters, probability = calculated probability)
```{r, message=FALSE}
municipalities <- 
  runoff %>%
  mutate(p_1 = number_clusters * turnout_municipality / turnout) %>%
  group_by(municipality, p_1) %>%
  summarise() %>%
  ungroup() %>%
  sample_n(size = number_clusters, weight = p_1) %>% 
  pull(municipality)

clusters <- 
  runoff %>% 
  filter(municipality %in% municipalities)

glimpse(clusters)
```

(10.) add a simple random sample as second stage to the cluster sample
```{r}
cl <- 
  clusters %>% 
  group_by(municipality) %>% 
  sample_n(sample_size / number_clusters) %>% 
  ungroup()

glimpse(cl)
```

