---
title: "Session6"
author: ""
date: "20/01/2021"
output: html_document
---

### Introduction
In today's session, we are working again with the survey on 5G. The focus here is on imputing missing values (i.e. fields where we do not know the answer). To do this, we follow the same steps as shown in Monday's video. 
In terms of content, we look at how the participants answer the question "Should Germany cooperate with the Chinese company Huawei in building the 5G network in Germany?".

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. loading packages
```{r}
pacman::p_load(haven, mice, mitools, naniar, readr, survey, tidyverse)
```

### 1. loading data
```{r}
df <- read_csv2("ZA7674_v1-0-0.csv")
```

### 2. preprocess data
```{r}
# defining NAs
na_strings <-
  c("weiß nicht, keine Angabe", "keine Angabe")

df_prep <-
  df %>%
  # select variables we want to use
  select(
    west_east = wo,
    sex = s1,
    age = s2,
    edu = s3,
    job = s4,
    party = s7,
    income = s9,
    f_g_heard = V6,
    f_g_pro = V7,
    huawei = V11,
    weight = gewicht
  ) %>%
  # replace NAs (defined strings)
  replace_with_na_all(condition = ~ .x %in% na_strings) %>% 
  # transform character variables to factor variables
  mutate_if(is.character, as.factor) %>% 
  # select only persons with knowledge on 5g
  filter(f_g_heard == "ja")
```

### 3. check data
use the map and the summary function to check if all variables are as expected

### 4. summary of missing values 
summarize missing values in the data numerically and/or graphically (at least one approach from the video)

### 5. predictor matrix
create a predictor matrix and set variables which shouldn't be used for imputation to 0

### 6. methods for imputation
using the mice function, create a vector containing the methods that will be used for imputation. Check if the methods are suitable and change the methods if necessary (an overview of the methods can be found in the documentation of the mice function).

### 7. imputation
use the mice function to impute the missing data (use the vector containing the methods and the predictor matrix you created before)

### 8. combining imputed data sets
use the complete function from the mice package and the imputationList function to create an imputed object

### 9. survey design object
create a survey design object for the imputed data and one for the unimputed data (just use weights)

### 10. estimate
calculate the mean for the huawei variable both for the imputed and unimputed data (optional: grouped by party)