---
title: "Roundup - Statistical Modelling"
author: "Lukas Weixler"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Useful terminology for any modelling discussion:
- Observation
- Variable
- Hypothesis Test
- Exogenous Variable
- Correlation
- Causal Effect
- Causal Path
- Identification
- Causal Collider
- Causal Fork
- Causal Chain of mediation
- Confounder
- Substantive Literature

- Go to [dagitty](http://www.dagitty.net/dags.html#) for causal identification of your model

```{r, message=F}
suppressWarnings(if (!require("pacman"))
  install.packages("pacman"))
pacman::p_load(ggdag, tidyverse)
```

```{r}
# importing our dataset
load('../SessionData/shoes.Rda')
```

Check out the dataset:
```{r}
glimpse(shoes)
```

Suspected relationship as a DAG in R
```{r}
coords <- list(
  x = c(
    shoe_size = 1,
    gender = 2,
    income = 3
  ),
  y = c(
    shoe_size = 1,
    gender = 2,
    income = 1
  )
)

dag <- dagify(income ~ shoe_size + gender,
              shoe_size ~ gender,
              coords = coords)

ggdag(dag,
      layout = "circle",
      text_size = 2.5,
      node_size = 18) +
  theme_dag()
```

Today, we'll try to explain reasons for high income.
Let's propose the following Null-Hypothesis: Shoe size doesn't have an effect on income.
```{r}
model_1 <- lm(income ~ shoe, data=shoes)

summary(model_1)
```

Check it out graphically:
```{r}

qplot(x=shoe, y=income, data=shoes)

qplot(x=shoe, y=income, data=shoes)+
  geom_smooth(method='lm')
```

So do we have to reject the H0 and recognize the effect?
Wait a minute and answer the following question: Are the two variables "shoe size" and "income" really exogenous?

Is there any confounding variable? 



Let's control for gender
```{r}
model2 <- lm(income ~ shoe + gender, data=shoes)
summary(model2)

```

The entire effect is actually due to gender, for which we 'll find tons of substantive literature.
```{r}
qplot(x=shoe, y=income, data=shoes, color = gender) +
  geom_jitter() +
  geom_smooth(method='lm')
```


