---
title: "ISM Tut - Session 3 - The Survey Package"
author: "Niklas Haffert"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparations
### Packages
* survey: analyzing survey data
* tidyverse: package for all tidy operations (dplyr, ggplot2, ...)
```{r, message=FALSE}
pacman::p_load(survey, tidyverse)
```

### Data
3 different samples we specified in the last session and the original population data frame
```{r load data}
# Simple Random Sampling
srs <-
  readRDS("../SessionData/usa_2016/srs.rds") %>%
  # specifying turnout for whole USA
  mutate(turnout_total = 136495547)

# Stratified Sampling
str <- readRDS("../SessionData/usa_2016/str.rds")

# Cluster Sampling
cl <- readRDS("../SessionData/usa_2016/cl.rds")
```

### New variables
creating number clusters variable for cluster sampled data (51 clusters as we have 51 "states")
```{r number clusters}
cl <-
  cl %>%
  mutate(number_clusters = 51)
```

# Survey Design Object
* svydesign object contains data and design information
* created from sample (data frame) 
* information about sampling design needs to be provided 

### Simple Random Sampling Object
* fpc: turnout (fpc = finite population correction)
* id: staging information (in srs only one stage, therefore 1)
```{r srs object}
srs_obj <-
  svydesign(data = srs,
            fpc = ~ turnout_total,
            id = ~ 1)
```

### Stratified Sampling Object 
* strata: indicates which column holds information about strata
* fpc changes from total population size to population size in strata
```{r str object}
str_obj <-
  svydesign(
    data = str,
    fpc = ~ turnout_state,
    id = ~ 1,
    strata = ~ state
  )
```

### Cluster Sampling Object (Multi Stage Sampling)
* id now reflects sampling stages: first stage clusters, second stage individuals
* fpc reflects population size in different stages: first stage number of clusters and second stage population size within the clusters
```{r clus object}
clus_obj <-
  svydesign(
    data = cl,
    id = ~ state + voter_id,
    fpc = ~ number_clusters + turnout_state
  )
```

# Estimates
With the svymean function the mean, based on the complex survey object, can be calculated (true values in the population: Clinton = 48.24%, Trump = 46.14%, Others = 5,62%)
```{r svymean srs}
svymean( ~ vote, design = srs_obj)
```

We can also calculate confidence intervals (95%)
```{r ci srs}
svymean(~ vote, design = srs_obj) %>%
  as.data.frame() %>%
  mutate(lower = mean - 1.96 * SE,
         upper = mean + 1.96 * SE)
```

With the svytotal function a population total is estimated (true values in the population: Clinton = 65,844,241, Trump = 62,979,031, Others = 7,672,275)
```{r svytotal str}
svytotal( ~ vote, design = str_obj)
```

Again CIs can be calculated
```{r ci str}
svytotal( ~ vote, design = str_obj) %>%
  as.data.frame() %>%
  mutate(lower = total - 1.96 * SE,
         upper = total + 1.96 * SE)
```

If you calculate the survey mean and the mean for the cluster sample without weighting according to the survey design, the difference becomes clear

mean by svymean:
```{r svymean cl}
svymean( ~ vote, design = clus_obj)
```

mean by standard mean function (without weights):
```{r mean cl}
cl %>%
  count(vote, name = "votes") %>%
  mutate(perc = votes / sum(votes))
```
