---
title: "ISM Tut - Session 3 - Extra"
author: "Niklas Haffert"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparations
### Packages
* data.table & R.utils loading data
* fastDummies: dummy variables from character variable
* survey: analyzing survey data
* tidyverse: package for all tidy operations (dplyr, ggplot2, ...)
```{r, message = FALSE}
pacman::p_load(data.table, R.utils, fastDummies, survey, tidyverse)
```

### Functions
Creating function to perform multiple srs draws from data
```{r}
srs_obj_fun <-
  function(data, sample_size) {
    data %>%
      # sampling individuals according to sample size
      sample_n(sample_size) %>%
      # adding fpc (size of population)
      mutate(fpc = nrow(data)) %>%
      # creating a survey design object with sampled data
      svydesign(data = .,
                fpc = ~ fpc,
                id = ~ 1)
  }

multiple_srs <- function(df, col, n_samples, sample_sizes) {
  # applying srs_obj_fun multiple times (given by n_samples)
  map(seq_len(n_samples),
      ~ srs_obj_fun(data = df, sample_size = sample_sizes)) %>%
    # calculating survey mean for each sample (col defines for which column)
    map(., ~ svymean(col, design = .)) %>%
    # creating one data frame with survey mean and standard errors for all samples
    map_df(~ as_tibble(.)) %>%
    rename(se = 2)
}
```

### Seed
Setting a seed allows to replicate random operations like sampling and get the same result
```{r set a seed}
set.seed(2020)
```

### Data
load the original population data frame
```{r load data}
votes <- fread("../SessionData/usa_2016/us_2016.gz") %>%
  as_tibble()
```

### new variable
dummy variables for three candidates
```{r dummy cols}
votes <- 
  votes %>%  
  # dummy (binary) variables for each candidate
  dummy_cols(select_columns = "vote",
             remove_selected_columns = TRUE) %>%
  rename(Clinton = 4,
         Other = 5,
         Trump = 6)
```

### Multiple Estimates
Here we use a function to draw 100 simple random samples of 1000 individuals each. The function returns the mean and standard error for each sample. On this basis we calculate the 95% confidence intervals. We estimate the vote share (popular vote) for Hillary Clinton.
```{r multiple samples}
true_val <- mean(votes$Clinton)

srs_df <-
  # sampling 
  multiple_srs(df = votes,
               col = ~ Clinton,
               n_samples = 100,
               sample_sizes = 1000) %>%
  # calculating 95% confidence intervals
  mutate(upper = mean + 1.96 * se,
         lower = mean - 1.96 * se) %>%
  # setting color according to if 95% confidence interval covers true value
  mutate(color = ifelse(true_val < upper &
                          true_val > lower, "green", "red")) %>% 
  arrange(mean) %>% 
  mutate(run = 1:nrow(.))

srs_df
```

Then we plot the point estimators (mean) and the confidence intervals. If the true value (solid black line) lies within the confidence interval, it is green, if not, it is red.
```{r plotting}
ggplot(srs_df, aes(x = mean, y = run, color = color)) +
  # true value
  geom_vline(aes(xintercept = true_val)) +
  # confidence intervals
  geom_linerange(aes(xmin = lower, xmax = upper)) +
  # point estimates
  geom_point() +
  scale_colour_identity() +
  xlab(label = "Estimate Percentage of Votes for Clinton") +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```
