---
title: "Session3"
author: "Niklas Haffert"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Sampling informations:

The srs data set contains a simple random sample. 
The str data set contains a stratified sample with municipalities as strata and number of individuals per strata according to number of voters per municipality (turnout_municipality)
The cl data set contains a cluster sample with voting station as cluster, 5 stations were sampled
The cl_multi data set contains a multistage cluster sample with municipalities as clusters and simple random samples within each sampled cluster. 

Number of municipalities and stations are stored in number_municipalities and number_stations. Turnout in turnout_municipality and turnout_station. The variable vote is indicating if an individual voted for Holland (1) or Sarkozy (0) in the second round of the 2012 French presidential election.

0. Settings for lonely PSU 
```{r}
options(survey.adjust.domain.lonely = TRUE)
options(survey.lonely.psu = "adjust")
```

1. Load the tidyverse and the survey package 
```{r}
pacman::p_load(survey, tidyverse)
```

2. Load the following data sets (srs, str, cl, cl_multi)
```{r}
srs <- readRDS("srs.rds")
str <- readRDS("str.rds")
cl <- readRDS("cl.rds")
cl_multi <- readRDS("cl_multi.rds")
```

3. Inspect the data
```{r}
glimpse(srs)
glimpse(str)
glimpse(cl)
glimpse(cl_multi)
```

4. Create survey design objects for each sample
```{r}
srs_obj <-
  svydesign(data = srs,
            fpc = ~ turnout,
            id = ~ 1)
```

```{r}
str_obj <-
  svydesign(
    data = str,
    fpc = ~ turnout_municipality,
    id = ~ 1,
    strata = ~ municipality
  )
```

```{r}
clus_obj <-
  svydesign(
    data = cl,
    id = ~ station,
    fpc = ~ number_stations
  )
```

```{r}
clus_multi_obj <-
  svydesign(
    data = cl_multi,
    id = ~ municipality + voter_id,
    fpc = ~ number_municipalities + turnout_municipality
  )
```

5. Use the svymean, svytotal or svyciprop functions for inference
```{r}
svymean(~vote, design = str_obj)
```

```{r}
svymean(~vote, design = srs_obj) %>% 
  as.data.frame() %>%
  rename(se = 2) %>% 
  mutate(lower = mean - 1.96 * se,
         upper = mean + 1.96 * se)
```

```{r}
svytotal(~vote, design = clus_obj)
```

```{r}
svymean(~vote, design = clus_multi_obj) %>% 
  as.data.frame() %>%
  rename(se = 2) %>% 
  mutate(lower = mean - 1.96 * se,
         upper = mean + 1.96 * se)

svyciprop(~vote, design = clus_multi_obj)
```

