---
title: "Session130121"
author: "Lukas Weixler"
date: "13 1 2021"
output: html_document
---

```{r}
pacman::p_load(tidyverse, survey)
```

### Generating today's fake dataset

Inspired by https://rpubs.com/mhanauer/268281

```{r}
set.seed(12345)

data <- tibble(.rows = 1000) %>% 
  mutate(income = sample(c(seq(0, 100000)), 1000, replace = TRUE),
         gender = sample(c("Male", "Female"), 1000, replace = TRUE),
         education = sample(c("Middleschool", "Highschool", "College", "PhD"), 1000, replace = TRUE)
         )
```

### Generating today's fake population distribution
```{r}
gender.dist <- tibble(gender = c("Male", "Female"),
                       Freq =  c(0.45, 0.55) * 80000000)
```


## Exercises:

#### Encode your fake dataset into a simple random sample surveydesign. Ignore the population distribution for now.

#### Post-Stratify your data on gender with the known population distribution


#### Rake your data on gender with the known population distribution


#### Generate an additional population distribution dataset on the education variable (you may choose quantities as you like.


#### Rake your data on gender and on education with the known population distributions


#### Compare the population estimates for your unweighted, your post-stratified, and your two raked surveydesigns.

